\documentclass[BIT, graphvis, a4paper]{usydthesis}

\usepackage[final]{pdfpages} 
\usepackage[toc]{appendix}
\usepackage[utf8]{inputenc}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{caption} 
\usepackage{color} 
\usepackage{chngpage}
\usepackage{float}
\usepackage{framed}
\usepackage{fullpage}
\usepackage{titletoc}
\usepackage{graphicx}
\usepackage[a4paper,margin=1.3in,headheight=20pt,footskip=.6in,headsep=.2in]{geometry}
\usepackage{hyperref} 
\usepackage{listings} 
\usepackage{subcaption} 
\usepackage[
	backend=biber,
	natbib=true,
	style=numeric,
	url=true, 
	doi=true,
	eprint=false
]{biblatex}
\usepackage{tabularx} 
\usepackage{tabu} 
\usepackage{tikz} 
\usepackage{url}
\usepackage{wrapfig} 
\usepackage{fancyhdr}
\pagestyle{fancy}

\renewcommand{\sectionmark}[1]{\markright{\thesection.\ #1}}

\fancyhead{} % clear all header fields
\fancyhead[LO,LE]{\leftmark}
\fancyhead[RO,RE]{\thesection}

\fancyfoot{}
\fancyfoot[RO,LE]{\thepage}
\fancyfoot[LO,RE]{Linus Karsai}

\renewcommand{\headrulewidth}{0.4pt}

\addbibresource{bibliography.bib}

% Related work title/citation
\newcommand{\relatedwork}[1]{
	\textit{\citetitle{#1} \textcite{#1}}
}


\usetikzlibrary{arrows}
\usetikzlibrary{shapes,snakes}
\usetikzlibrary{positioning,calc}
\DeclareUnicodeCharacter{00A0}{ }
\newcommand*\rot{\rotatebox{90}}
\makeatletter
\newcommand{\@chapapp}{\relax}%
\makeatother
\graphicspath{ {images/} }
\setlength{\parskip}{\baselineskip}%
\setlength{\parindent}{0pt}%
% graph settings
\definecolor{nodeyellow}{RGB}{255,252,135}
\definecolor{nodeblue}{RGB}{179,255,255}
\tikzset{
	every node/.style={font=\sffamily\scriptsize}
}
\tikzset{
	group/.style={font=\sffamily,fill=nodeblue}
}
\tikzset{
	->,>=stealth',shorten >=1pt,auto,node distance=3cm,
	thick,main node/.style={ellipse,fill=nodeyellow,draw,font=\sffamily\normalsize}
}

% give captions a bit of space.
\captionsetup[sub]{margin=5pt, skip=10pt}

% listings style for prov code
\lstdefinestyle{provn}{
	breaklines=true,
	showstringspaces=false,
	basicstyle=\footnotesize\ttfamily,
}
\lstdefinestyle{basic}{
	breaklines=true,
	showstringspaces=false,
	basicstyle=\footnotesize\ttfamily,
	keywordstyle=\color{black},
	commentstyle=\color{black},
	stringstyle=\color{black},
}


\title{Clustering Provenance}

\begin{document}
\author{Linus Karsai}

\pagenumbering{roman}
\maketitle
\chapter*{Contents}

\startcontents[sections]
\printcontents[sections]{l}{0}{\setcounter{tocdepth}{1}}

\include{abstract} % abstract of thesis

% reset the page numbering and change to arabic numbers
\newpage\setcounter{page}{1}\pagenumbering{arabic}

\include{introduction} % intro to pro etc.

\include{litreview} % literature review

\section{Contributions}

The related work above shows that a growing amount of research is been conducted in the field of provenance. There's already ways to capture provenance (Section~\ref{sec:provenance_acquisition}) as well as ways to store it once its captured (Section~\ref{sec:provenance_storage}). Research has also been conducted into how to then secure this information so that access can be controlled nor can provenance be altered arbitrarily (Section~\ref{sec:provenance_security}). However there a gap in the research when it comes to creating interfaces that allows users to cluster effectively.

As mentioned in Section~\ref{sec:provenance_display}, there is already some existing applications that can be used to visualise provenance graphs. However most of them lack clustering and in particular clustering from a usability perspective. This leaves the following problem unsolved: \textit{Provenance graphs are too complex for regular users to understand.} 

In the upcoming sections of my thesis I outline and go into detail about my three main contributions to the field:

\begin{itemize}
	\item The design of an interface that enables users to cluster effectively.
	\item The creation of a working prototype that implements the aforementioned design.
	\item An evaluation of the design and prototype: covering learnability, ease of use, efficiency, recovery from errors, user attitude.
\end{itemize}


\include{challenges} % challenges of clustering

\include{interface} % explanation of interface

\include{technical-decisions} % technical implementation decisions

\include{usability-plan} % plan of question coding etc.

\section{Results} % results

\include{conclusion}

\printbibliography

\include{appendix}

\end{document}
